Deep Reinforcement Learning for Conservation Decisions

### 2021-08-12 manuscript/manuscript.Rmd
Small changes; paragraphs on behav. etc. otw to the discussion

<!-- Environmental problems, ML -->
Advances in both available data and computing power are opening the door for machine learning (ML) to play a greater role in addressing some of our planet’s most pressing environmental problems. 
But will ML approaches really help us tackle our most pressing environmental problems? From the growing frequency and intensity of wildfire [@wildfire], to over-exploited fisheries [@Worm2006] and declining biodiversity [@biodiversity], to emergent zoonotic pandemics [@covid], the diversity and scope of environmental problems are unprecedented.
Applications of ML in ecology have to-date illustrated the promise of two methods: *supervised learning* [@joseph_neural_2020] and *unsupervised learning* [@valletta_applications_2017]. 
However, the fields of ecology and conservation science have largely overlooked the third and possibly most promising approach in the ML triad: *reinforcement learning* (RL).
Three features distinguish RL from other ML methods in ways that are particularly well suited to addressing issues of global ecological change:

@@

Despite relevance to decision making under uncertainty that could make RL uniquely well suited for ecological and conservation problems, RL has not been widely applied across these fields except in behavioral ecology [@yamada_evolution_2020, @wang_reinforcement_2020].
To date, the problems considered by RL research have largely been drawn from examples in robotic movement and games like Go and Starcraft [@openai_learning_2019; @alphazero; @alphastar].
Complex environmental problems share many similarities to these tasks and games: the need to plan many moves ahead given a large number of possible outcomes, to account for uncertainty and to respond with contingency to the unexpected.
RL agents typically develop strategies by interacting with simulators, a practice that should not be unsettling to ecologists since learning from simulators is common across ecology.
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or climate change policy models [@dice] are already used to explore scenarios and inform ecosystem management.
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of ecological models; the so called "curse of dimensionality" makes problems with a large number of states or actions intractable by conventional methods [@Possingham2006; @Marescot2013; @ferrer-mestres_k-n-momdps_2021; @chades_primer_2021].
Neural-network-based RL techniques, referred to as *deep RL*, have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.

@@

How would an RL agent be applied to empirical data?
In principle, this is straightforward.
After we train an agent on a simulation environment that approximates the fishery of interest, we can query the policy of the agent to find a quota for the observed stock.
To illustrate this, we examine the quota that would be recommended by our newly trained RL agent, above, against historical harvest levels of Argentine hake based on stock assessments from 1986 - 2014 [@ramlegacy, see Appendix D].
Hake stocks showed a marked decline throughout this period, while harvests decreased only in proportion [Fig 3].
In contrast, our RL agent would have recommended significantly lower quotas over most of the same interval, including the closure of the fishery as stocks were sufficiently depleted.
While we have no way of knowing for sure if the RL quotas would have led to recovery, let alone an optimal harvest rate, the contrast between those recommended quotas and the historical catch is notable.

@@

<!-- Model free is great bc we don't know the true models -->
The rapidly expanding class of model-free RL algorithms is particularly appealing given the ubiquitous presence of model uncertainty in ecological dynamics.
Rarely do we know the underlying functional forms for ecological processes.
Methods which must first assume something about the structure or functional form of a process before estimating the corresponding parameter can only ever be as good as those structural assumptions.
Frequently available ecological data is insufficient to distinguish between possible alternative models [@knape_are_2012], or the correct model may be non-identifiable with any amount of data.
Model-free RL approaches offer a powerful solution for this thorny issue.
Model-free algorithms have proven successful at learning effective policies even when the underlying model is difficult or impossible to learn [@Pong2020], as long as simulations of possible mechanisms are available.
Yet, model-free algorithms tend to have poor sample efficiency [@sac] which may restrain them from being insightful on simulators that are slow to run.

>>>>

<!-- Environmental problems, ML -->
Advances in both available data and computing power are opening the door for machine learning (ML) to play a greater role in addressing some of our planet’s most pressing environmental problems. 
But will ML approaches really help us tackle our most pressing environmental problems? From the growing frequency and intensity of wildfire [@wildfire], to over-exploited fisheries [@Worm2006] and declining biodiversity [@biodiversity], to emergent zoonotic pandemics [@covid], the diversity and scope of environmental problems are unprecedented.
Applications of ML in ecology have to-date illustrated the promise of two methods: *supervised learning* [@joseph_neural_2020] and *unsupervised learning* [@valletta_applications_2017]. 
However, the fields of ecology and conservation have largely overlooked the third and possibly most promising approach in the ML triad: *reinforcement learning* (RL).
Three features distinguish RL from other ML methods in ways that are particularly well suited to addressing issues of global ecological change:

@@

Despite relevance to decision making under uncertainty that could make RL uniquely well suited for ecological conservation, RL has not been widely applied across this field.
To date, the problems considered by RL research have largely been drawn from examples in robotic movement and games like Go and Starcraft [@openai_learning_2019; @alphazero; @alphastar].
Complex environmental problems share many similarities to these tasks and games: the need to plan many moves ahead given a large number of possible outcomes, to account for uncertainty and to respond with contingency to the unexpected.
RL agents typically develop strategies by interacting with simulators, a practice that should not be unsettling to ecologists since learning from simulators is common across ecology.
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or climate change policy models [@dice] are already used to explore scenarios and inform ecosystem management.
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of ecological models; the so called "curse of dimensionality" makes problems with a large number of states or actions intractable by conventional methods [@Possingham2006; @Marescot2013; @ferrer-mestres_k-n-momdps_2021; @chades_primer_2021].
Neural-network-based RL techniques, referred to as *deep RL*, have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.

@@

How could an RL agent be applied to empirical data?
One solution would be to train an agent on a simulation environment that approximates the fishery of interest then query the policy of the agent to find a quota for the observed stock.
To illustrate this, we examine the quota that would be recommended by our newly trained RL agent, above, against historical harvest levels of Argentine hake based on stock assessments from 1986 - 2014 [@ramlegacy, see Appendix D].
Hake stocks showed a marked decline throughout this period, while harvests decreased only in proportion [Fig 3].
In contrast, our RL agent would have recommended significantly lower quotas over most of the same interval, including the closure of the fishery as stocks were sufficiently depleted.
While we have no way of knowing for sure if the RL quotas would have led to recovery, let alone an optimal harvest rate, the contrast between those recommended quotas and the historical catch is notable.

@@

<!-- Model free is great bc we don't know the true models -->
The rapidly expanding class of model-free RL algorithms is particularly appealing given the ubiquitous presence of model uncertainty in ecological dynamics.
Rarely do we know the underlying functional forms for ecological processes.
Methods which must first assume something about the structure or functional form of a process before estimating the corresponding parameter can only ever be as good as those structural assumptions.
Frequently available ecological data is insufficient to distinguish between possible alternative models [@knape_are_2012], or the correct model may be non-identifiable with any amount of data.
Model-free RL approaches offer a powerful solution for this thorny issue.
Model-free algorithms have proven successful at learning effective policies even when the underlying model is difficult or impossible to learn [@Pong2020], as long as simulations of possible mechanisms are available.
Yet, current model-free algorithms tend to have poor sample efficiency [@sac] which may restrain them from being insightful on simulators that are slow to run.
